---
title: "H17w (GSE112570) Data Analysis"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

This notebook generates the figures for paper XXXX

```{r,message=F,warning=F}
set.seed(408)
library(Seurat)
library(dplyr)
library(monocle)
library(WGCNA)
library(igraph)
library(mclust)

# Loading R objects for fast knitting
load("_robjs/H17w/seurat-human17w-nephro.Robj")
load("_robjs/H17w/seurat-human17w.Robj")
load("_robjs/H17w/mclust-human17w-nephro.Robj")
load("_robjs/H17w/monocle-human17w.Robj")
load("_robjs/H17w/WGCNA-human17w-nephro.Robj")

# Loads sources of helper functions
source("R/seurat.R")
source("R/monocle.R")
source("R/wgcna.R")
source("R/clustering.R")

```

# Analysis of the whole dataset

 Load sparse matrices 
```{r,eval=F}
df.h17w.k1 <- Read10X("counts/10x/Human/H17w/K1")
df.h17w.k2 <- Read10X("counts/10x/Human/H17w/K2")
```

 Append K1 and K2 to colnames so we know which kidney it comes from, then join datasets
```{r,eval=F}
colnames(df.h17w.k1) <- paste0("K1_", colnames(df.h17w.k1))
colnames(df.h17w.k2) <- paste0("K2_", colnames(df.h17w.k2))
df.h17w <- cbind(df.h17w.k1, df.h17w.k2)
```

 Creates Seurat Object from dataframe
```{r,eval=F, warning=F}
srt.h17w <- CreateSeuratObject(raw.data = df.h17w, min.cells = 0, min.genes = 0, scale.factor = 1e4, do.center=F, project = "H17w", normalization.method = "LogNormalize")
```

 Calculates Saturation and mitochondrial pct
```{r,eval=F}
saturation <- 1 - apply(df.h17w == 1, 2, sum)/apply(df.h17w, 2, sum)
pct.mito <- apply(df.h17w[grepl("^MT-", rownames(df.h17w)),], 2, sum)/apply(df.h17w, 2, sum)
```


 QC on different metrics: Number of genes, saturation, mitochondrial %
```{r,eval=F}
hist(saturation,breaks = 100)
hist(pct.mito,breaks = 100)
hist(srt.h17w@meta.data$nGene, breaks = 100)
srt.h17w <- AddMetaData(srt.h17w, saturation, "saturation")
srt.h17w <- AddMetaData(srt.h17w, pct.mito, "pct.mito")

# Subsets data like in the methods
head(srt.h17w@meta.data)
srt.h17w <- SubsetData(srt.h17w, subset.name = "saturation", accept.low = .7)
srt.h17w <- SubsetData(srt.h17w, subset.name = "nGene", accept.low = 1000)
srt.h17w <- SubsetData(srt.h17w, subset.name = "pct.mito", accept.high = .05)
```

 Finds variable genes
```{r,eval=F}
srt.h17w <- ScaleData(srt.h17w, vars.to.regress = c("nGene", "nUMI", "saturation", "orig.ident"))
srt.h17w <- FindVariableGenes(srt.h17w, x.low.cutoff = .05, x.high.cutoff = Inf, y.cutoff = .5, y.high.cutoff = Inf)
length(srt.h17w@var.genes)
srt.h17w <- RunPCA(srt.h17w, pc.genes = srt.h17w@var.genes, pcs.compute = 40)
```

 Runs JackStraw to find significant PCs
```{r,eval=F}
srt.h17w <- JackStraw(srt.h17w, num.pc = 40, do.print=T)
JackStrawPlot(srt.h17w, PCs = 1:40, nCol = 6)
```

 Runs TSNE and Seurat clustering on the whole dataset
```{r, message=F,eval=F}
srt.h17w <- RunTSNE(srt.h17w, dims.use = 1:39, do.fast = T, verbose=T, perplexity = 2000)
srt.h17w <- FindClusters(srt.h17w, dims.use = 1:39)
srt.h17w <- BuildClusterTree(srt.h17w, do.reorder=T, reorder.numeric=T)
```

# Supplementary Fig 4A
```{r}
TSNEPlot(srt.h17w, group.by = "orig.ident", do.label=T)
TSNEPlot(srt.h17w, do.label=T)
```

# Analysis of the nephro lineage

 Creates a subset with just the nephro lineage
```{r,eval=F}
srt.h17w.nephro <- SubsetData(srt.h17w, ident.use = c(11,13,14,15,16))
```

 Finds variable genes in the subset of cells
```{r,eval=F}
srt.h17w.nephro <- FindVariableGenes(srt.h17w.nephro, x.low.cutoff = .05, x.high.cutoff = Inf, y.cutoff = 1, y.high.cutoff = Inf)
```

 Runs PCA
```{r,eval=F}
srt.h17w.nephro <- RunPCA(srt.h17w.nephro, pc.genes = srt.h17w.nephro@var.genes, pcs.compute = 50)
```

 Runs JackStraw
```{r,eval=F}
srt.h17w.nephro <- JackStraw(srt.h17w.nephro, num.pc = 50, do.print = T) # 19 PCs with p < .05
```

 Run TSNE with 19 PCs
```{r,eval=F}
srt.h17w.nephro <- RunTSNE(srt.h17w.nephro, dims.use = 1:19, do.fast = T)
```

 Clusters with GMM and assigns it to the seurat object
```{r,eval=F}
res.h17w <- Mclust(srt.h17w.nephro@dr$pca@cell.embeddings[, 1:19], G = 1:50)
srt.h17w.nephro@ident <- factor(res.h17w$classification)
```

 Figure 2A
```{r}
TSNEPlot(srt.h17w.nephro, do.label=T)
```

 Differential Expression
```{r,eval=F}
markers.h17w.nephro <- FindAllMarkers(srt.h17w.nephro, logfc.threshold = 0, min.pct = 0, only.pos = T)
```

 Figure 2C
```{r}
bds <- bd.seurat(srt.h17w.nephro, PCs = 1:19)
plot.pairwise.bd(bds)
```

 Supplementary Fig 3B
```{r}
plot.bd.mst(bds)
```

# Monocle Analysis
 Turns Monocle into Seurat object
```{r,eval=F}
mon.17w.nephro <- monocle.from.seurat(srt.h17w.nephro)
mon.17w.nephro.p2345 <- mon.sub(mon.17w.nephro, c(2,3,4,5))
```

 Figure 2D
```{r}
plot_cell_trajectory(mon.17w.nephro, color_by="ident")
plot_cell_trajectory(mon.17w.nephro.p2345, color_by="ident")
```

# Module analysis

```{r,eval=F}
datExpr <- GetDatExpr(srt.h17w.nephro, markers.h17w.nephro)
```

```{r}
sft <- getSft(datExpr)
sft # We pick the power = 3 to maximize truncated Rsq. 
```

 Finds modules 
```{r,eval=F,message=F}
net <- Modules(datExpr, p = 3)
```

 Figure 3A (top)
```{r,message=F}
BoxMEByCluster(srt.h17w.nephro, datExpr, net, "MEturquoise")
BoxMEByCluster(srt.h17w.nephro, datExpr, net, "MEyellow")
BoxMEByCluster(srt.h17w.nephro, datExpr, net, "MEblue")
BoxMEByCluster(srt.h17w.nephro, datExpr, net, "MEblack")
BoxMEByCluster(srt.h17w.nephro, datExpr, net, "MEgreenyellow")
BoxMEByCluster(srt.h17w.nephro, datExpr, net, "MEtan")
BoxMEByCluster(srt.h17w.nephro, datExpr, net, "MEmagenta")
BoxMEByCluster(srt.h17w.nephro, datExpr, net, "MEbrown")
BoxMEByCluster(srt.h17w.nephro, datExpr, net, "MEred")
BoxMEByCluster(srt.h17w.nephro, datExpr, net, "MEgreen")
BoxMEByCluster(srt.h17w.nephro, datExpr, net, "MEpurple")
BoxMEByCluster(srt.h17w.nephro, datExpr, net, "MEpink")
```

 Add Eigengene values to Seurat
```{r,eval=F}
srt.h17w.nephro <- AddMEsToMetaData(srt.h17w.nephro, net)
```

 Figure 3A (top)
```{r,message=F}
TSNEPlotFeature(srt.h17w.nephro, "MEturquoise")
TSNEPlotFeature(srt.h17w.nephro, "MEyellow")
TSNEPlotFeature(srt.h17w.nephro, "MEblue")
TSNEPlotFeature(srt.h17w.nephro, "MEblack")
TSNEPlotFeature(srt.h17w.nephro, "MEgreenyellow")
TSNEPlotFeature(srt.h17w.nephro, "MEtan")
TSNEPlotFeature(srt.h17w.nephro, "MEmagenta")
TSNEPlotFeature(srt.h17w.nephro, "MEbrown")
TSNEPlotFeature(srt.h17w.nephro, "MEred")
TSNEPlotFeature(srt.h17w.nephro, "MEgreen")
TSNEPlotFeature(srt.h17w.nephro, "MEpurple")
TSNEPlotFeature(srt.h17w.nephro, "MEpink")
```

 Pseudotime-Eigengene curve-fitting
```{r}
scores <- ModuleScore(datExpr, power = 3, net = net, mon = mon.17w.nephro.p2345)
```

 Figure 3B
```{r}
PlotScores(scores, net)
```


